\documentclass[12pt]{exam}

\newcommand{\ds}{\ensuremath{\displaystyle}}

% \printanswers

\begin{document}

\begin{center}
\fbox{\fbox{\parbox{5.5in}{\centering
Calculus II - Fall 2014 - Mr. Clontz - PRACTICE Midterm
}}}
\end{center}
\vspace{0.1in}
\makebox[\textwidth]{Name:\enspace\hrulefill\space 9am / 10am}

\vspace{12pt}

\noindent About your midterm and this practice midterm:
\begin{itemize}
  \item
    The first few questions will be multiple-choice questions
    covering basic definitions, theorems, and other concepts in sections
    11.1-11.10, 7.1, and 7.2.
    These will total 10 points of the midterm, and are not covered
    by the practice midterm.
  \item
    The other 90 points of the midterm are based on 9 of the questions
    asked on this practice midterm. These questions will require full solutions
    and will be given partial credit based on the rubric for each question
    in this practice midterm. The rubric will not be given on the actual
    midterm.
  \item
    A review for the midterm will be held on Friday, October 3 during lecture.
    Students will receive presentation credit for solving problems from the
    practice midterm.
  \item
    The midterm will take place during lecture on Monday, October 6. Up to
    ten bonus points will be awarded for turning in full solutions to questions
    asked on this practice midterm before taking the midterm.
\end{itemize}

\newpage

\begin{questions}

\setcounter{question}{0}
\question[10]
Find the radius of convergence for the series
$\ds \sum_{n=0}^{\infty} \frac{2n\left(x-1\right)^n}{5^{n+1}}$

\begin{checkboxes}
\choice $R=2$
\choice $R=\frac{1}{2}$
\CorrectChoice $R=5$
\choice $R=\frac{1}{5}$
\choice None of these.
\end{checkboxes}

\vfill

\question[10]
Find a power series representation for $\ln\left(1+x\right)$.

\begin{checkboxes}
\choice $\ds\sum_{n=0}^\infty \frac{(-1)^nx^{2n}}{(2n)!}$
\choice $\ds\sum_{n=0}^\infty \frac{x^n}{n!}$
\CorrectChoice $\ds\sum_{n=0}^\infty \frac{x^{n+1}}{n+1}$
\choice $\ds\sum_{n=0}^\infty (-1)^nx^{2n+1}$
\choice A different series. (Write your answer for credit.)
\end{checkboxes}

\vfill

\question[5]
Find the Maclaurin Series for $f(x) = e^x$.

\begin{checkboxes}
\choice $\ds\sum_{n=0}^\infty \frac{(-1)^nx^{2n}}{(2n)!}$
\CorrectChoice $\ds\sum_{n=0}^\infty \frac{x^n}{n!}$
\choice $\ds\sum_{n=0}^\infty \frac{x^{n+1}}{n+1}$
\choice $\ds\sum_{n=0}^\infty (-1)^nx^{2n+1}$
\choice A different series. (Write your answer for credit.)
\end{checkboxes}

\vfill

\end{questions}


\end{document}